pdp: An R Package for Constructing Partial Dependence Plots

### 2016-12-15 manuscript/greenwell.tex
tweaks based on review comments

\abstract{
Complex nonparametric models---like neural networks, random forests, and support vector machines---are more common than ever in predictive analytics, especially when dealing with large observational databases that don't adhere to the strict assumptions imposed by traditional statistical techniques (e.g., multiple linear regression which assumes linearity and normality). Unfortunately, it can be challenging to understand the results of such models and explain them to management. Partial dependence plots offer a simple solution. Partial dependence plots are low-dimensional graphical renderings of the prediction function $\widehat{f}\left(\boldsymbol{x}\right)$ so that the relationship between the outcome and predictors of interest can be
more easily understood. These plots are especially useful in explaining the output from black box models. In this paper, we introduce \pkg{pdp}, a general R package for constructing partial dependence plots.
}

@@

>>>>

\abstract{
Complex nonparametric models---like neural networks, random forests, and support vector machines---are more common than ever in predictive analytics, especially when dealing with large observational databases that don't adhere to the strict assumptions imposed by traditional statistical techniques (e.g., multiple linear regression which assumes linearity, homoscedasticity, and normality). Unfortunately, it can be challenging to understand the results of such models and explain them to management. Partial dependence plots offer a simple solution. Partial dependence plots are low-dimensional graphical renderings of the prediction function $\widehat{f}\left(\boldsymbol{x}\right)$ so that the relationship between the outcome and predictors of interest can be
more easily understood. These plots are especially useful in explaining the output from black box models. In this paper, we introduce \pkg{pdp}, a general R package for constructing partial dependence plots.
}

@@

\citet{harrison-1978-hedonic} were among the first to analyze the well-known Boston housing data. One of their goals was to find a housing value equation using data on median housing values from 506 census tracts in the suburbs of Boston from the 1970 census. The prediction equation from one of there regression analyses is reproduced in Equation~\eqref{eqn:boston}. \citet{harrison-1978-hedonic} took great care and time to find this model which obtained an $R^2$ of roughly 0.81. Part of the payoff for there efforts was a reaonable fitting model with an interpretable equation. For example, the equation tells us that an increase in median home value is associated with an increase in the average number of rooms per dwelling.
\begin{equation}
\label{eqn:boston}
\begin{split}
\widehat{\log\left(MV\right)} &= 9.76 + 0.0063 RM^2 + 8.98\times10^{-5} AGE - 0.19\log\left(DIS\right) + 0.096\log\left(RAD\right) \\
  &\quad - 4.20\times10^{-4} TAX - 0.031 PTRATIO + 0.36\left(B - 0.63\right)^2 - 0.37\log\left(LSTAT\right) \\
  &\quad - 0.012 CRIM + 8.03\time10^{-5} ZN + 2.41\times10^{-4} INDUS + 0.088 CHAS - 0.0064 NOX^2
\end{split}
\end{equation}

Many modern nonparametric models, however, can fit the data automatically in seconds---typically with higher accuracy. The downfall, however, is some loss of interpretation; for example, there is usually no prediction formula like in Equation~\eqref{eqn:boston}. Fortunately, methods like boosting and random forest have a natural way to assign variable importance scroes to the predictors which helps gives some insight into the data.
